### 2008-01-20 properties.tex
Itemize-based tabels for properties & models.

\noindent
\textbf{Network:}
\begin{itemize}
  \item collection of nodes
  \item collection of flows
  \item distribution of flow sizes
  \item distribution of intervals between flow start times
  \item distribution of packet sizes
  \item distribution of inter-packet intervals
\end{itemize}
\noindent
\textbf{Node:}
\begin{itemize}
  \item collection of source flows
  \item collection of destination flows
  \item session start time
  \item distribution of flow sizes
  \item distribution of intervals between flow start times
  \item distribution of packet sizes
  \item distribution of inter-packet intervals
\end{itemize}
\noindent
\textbf{Flow:}
\begin{itemize}
  \item source
  \item destination
  \item start time
  \item size (byte count)
  \item distribution of packet sizes
  \item distribution of inter-packet intervals
\end{itemize}
\noindent
\textbf{Packet:}
\begin{itemize}
  \item send time
  \item payload size (bytes)
\end{itemize}