### 2008-01-20 models.tex
Added subtle bullets.

\begin{tabular}{|c|c|p{4.3in}|}
\multicolumn{1}{c}{\textbf{Level}} &
\multicolumn{1}{c}{\textbf{Model}} &
\multicolumn{1}{c}{\textbf{Description}} \\[2pt]
\hline
\multirow{8}{*}[2.5pt]{Node}
& \multirow{1}{*}[-0.05em]{Individual} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each node corresponds to a specific trace node with associated trace flows \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned from trace \\
\raisebox{1.5pt}{$\centerdot$} overrides all distributions with per-node versions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Nonparametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes by sampling \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned from pool of start times
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Parametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes according to BiPareto model \\
\raisebox{1.5pt}{$\centerdot$} session start time randomly chosen from viable range
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Uniform} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all nodes have average number of generic flows \\
\raisebox{1.5pt}{$\centerdot$} session start times uniformly cover viable range
\vspace{2pt}
\end{minipage} \\
\hline
\multirow{11}{*}[2.5pt]{Flow}
& \multirow{1}{*}[-0.05em]{Individual} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each flow corresponds to a specific trace flow with an associated packet trace \\
\raisebox{1.5pt}{$\centerdot$} flow size \& start time taken from trace \\
\raisebox{1.5pt}{$\centerdot$} overrides distributions of size/interval packet behaviors with flow-specific versions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Nonparametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} flow size is sampled from inherited distribution of flow sizes \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous session flow sampled from inherited distribution \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Parametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} flow size is sampled from BiPareto model fitted to inherited distribution \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous session flow sampled Lognormal fitted to inherited \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Uniform} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all flows have size equal to inherited mean \\
\raisebox{1.5pt}{$\centerdot$} average inter-flow interval used between all flows \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\hline
\multirow{5}{*}[2.5pt]{Packet}
& \multirow{1}{*}[-0.05em]{Individual} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each packet corresponds to a specific trace packet
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Nonparametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals are sampled from inherited distributions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Parametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals are drawn from inherited parametric distributions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Uniform} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all packet sizes and inter-packet intervals are equal to inherited mean
\vspace{2pt}
\end{minipage} \\
\hline
\end{tabular}