### 2008-01-21 models.tex
Various edits:
  - models.rb:
  	- removed unnecessary #leaves method from Object, Array and Hash
  	- turn "*" at end of line into footnote mark
  - models.yaml:
    - regularized wording of Parametric model entries
    - added footnote markers ("*") at end of fitted Parametric entries
  - paper.tex:
    - changed title
    - work on introduction
    - added footnote text to models table (parametrics fitted with MLE)
  - references.bib:
    - added back Royer01 entry that got lost in the shuffle

\begin{tabular}{|c|c|p{4.3in}|}
\multicolumn{1}{c}{\textbf{Level}} &
\multicolumn{1}{c}{\textbf{Model}} &
\multicolumn{1}{c}{\textbf{Overridden and inherited properties}} \\[2pt]
\hline
\multirow{8}{*}[2.5pt]{Node}
& \multirow{1}{*}[-0.05em]{Individual} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each node corresponds to a specific trace node with associated trace flows \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned from trace \\
\raisebox{1.5pt}{$\centerdot$} overrides all distributions with node-specific versions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Nonparametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes by sampling [TODO---more detail] \\
\raisebox{1.5pt}{$\centerdot$} session start time randomly assigned from pool of trace session start times
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Parametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes according to fitted BiPareto model\raisebox{1pt}{\footnotesize*} \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned according to Poisson arrival model\raisebox{1pt}{\footnotesize*}
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Uniform} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all nodes have average number of generic flows \\
\raisebox{1.5pt}{$\centerdot$} session start times uniformly cover viable range
\vspace{2pt}
\end{minipage} \\
\hline
\multirow{11}{*}[2.5pt]{Flow}
& \multirow{1}{*}[-0.05em]{Individual} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each flow corresponds to a specific trace flow with an associated packet trace \\
\raisebox{1.5pt}{$\centerdot$} flow size and start time taken from trace \\
\raisebox{1.5pt}{$\centerdot$} overrides distributions of size/interval packet behaviors with flow-specific versions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Nonparametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} flow size is sampled from inherited distribution of flow sizes \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous session flow sampled from inherited distribution \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Parametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} flow size is sampled from fitted BiPareto model\raisebox{1pt}{\footnotesize*} \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous flow in session sampled from fitted Lognormal model\raisebox{1pt}{\footnotesize*} \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Uniform} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all flows have size equal to inherited mean \\
\raisebox{1.5pt}{$\centerdot$} average inter-flow interval used between all flows \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\hline
\multirow{5}{*}[2.5pt]{Packet}
& \multirow{1}{*}[-0.05em]{Individual} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each packet corresponds to a specific trace packet
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Nonparametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals sampled from inherited distributions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Parametric} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals drawn from fitted Exponential distributions\raisebox{1pt}{\footnotesize*}
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{Uniform} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all packet sizes and inter-packet intervals are equal to inherited mean
\vspace{2pt}
\end{minipage} \\
\hline
\end{tabular}