### 2008-01-21 models.tex
Redone intro; added paragraph to related work; etc.
In the process of overhauling the Methodology section.

\begin{tabular}{|c|c|p{4.3in}|}
\multicolumn{1}{c}{\textbf{Level}} &
\multicolumn{1}{c}{\textbf{Sub-model}} &
\multicolumn{1}{c}{\textbf{Overridden and inherited properties}} \\[2pt]
\hline
\multirow{8}{*}[2.5pt]{\class{Node}}
& \multirow{1}{*}[-0.05em]{\submodel{Trace}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each node corresponds to a specific trace node with associated trace flows \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned from trace \\
\raisebox{1.5pt}{$\centerdot$} overrides all distributions with node-specific versions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Nonparametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes by sampling [TODO---more detail] \\
\raisebox{1.5pt}{$\centerdot$} session start time randomly assigned from pool of trace session start times
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Parametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes according to fitted BiPareto model\raisebox{1.5pt}{\scriptsize*} \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned according to Poisson arrival model\raisebox{1.5pt}{\scriptsize*}
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Uniform}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all nodes have average number of generic flows \\
\raisebox{1.5pt}{$\centerdot$} session start times uniformly cover viable range
\vspace{2pt}
\end{minipage} \\
\hline
\multirow{11}{*}[2.5pt]{\class{Flow}}
& \multirow{1}{*}[-0.05em]{\submodel{Trace}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each flow corresponds to a specific trace flow with an associated packet trace \\
\raisebox{1.5pt}{$\centerdot$} flow size and start time taken from trace \\
\raisebox{1.5pt}{$\centerdot$} overrides distributions of size/interval packet behaviors with flow-specific versions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Nonparametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} flow size is sampled from inherited distribution of flow sizes \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous session flow sampled from inherited distribution \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Parametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} flow size is sampled from fitted BiPareto model\raisebox{1.5pt}{\scriptsize*} \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous flow in session sampled from fitted Lognormal model\raisebox{1.5pt}{\scriptsize*} \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Uniform}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all flows have size equal to inherited mean \\
\raisebox{1.5pt}{$\centerdot$} average inter-flow interval used between all flows \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node
\vspace{2pt}
\end{minipage} \\
\hline
\multirow{5}{*}[2.5pt]{\class{Packet}}
& \multirow{1}{*}[-0.05em]{\submodel{Trace}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each packet corresponds to a specific trace packet
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Nonparametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals sampled from inherited distributions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Parametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals drawn from fitted Exponential distributions\raisebox{1.5pt}{\scriptsize*}
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\submodel{Uniform}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all packet sizes and inter-packet intervals are equal to inherited mean
\vspace{2pt}
\end{minipage} \\
\hline
\end{tabular}