### 2008-01-21 properties.tex

\textbf{\submodel{Network:}}\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~collection of nodes\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~collection of flows\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of flow sizes\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of flow-source counts\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of flow-destination counts\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of intervals between flow start times\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of packet sizes\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of inter-packet intervals\\
\textbf{\submodel{Node:}}\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~session start time\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~collection of source flows\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~collection of destination flows\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of flow sizes\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of intervals between flow start times\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of packet sizes\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of inter-packet intervals\\
\textbf{\submodel{Flow:}}\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~start time\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~source node\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~destination node\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~stream of packets\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of packet sizes\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~distribution of inter-packet intervals\\
\textbf{\submodel{Packet:}}\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~send time\\
\hspace*{1ex}\raisebox{1.5pt}{$\centerdot$}~payload size (bytes)\\