### 2008-01-21 models.tex
Fixed up formatting of TOM classes.

\begin{tabular}{|c|c|p{4.3in}|}
\multicolumn{1}{c}{\textbf{Class}} &
\multicolumn{1}{c}{\textbf{Subclass}} &
\multicolumn{1}{c}{\textbf{Overridden and inherited properties}} \\[2pt]
\hline
\multirow{8}{*}[2.5pt]{\class{Node}}
& \multirow{1}{*}[-0.05em]{\class{Trace}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} session start time assigned from trace \\
\raisebox{1.5pt}{$\centerdot$} each node corresponds to a specific trace node with associated trace flows \\
\raisebox{1.5pt}{$\centerdot$} overrides all distributions with node-specific versions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Nonparametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} session start time randomly assigned from pool of trace session start times \\
\raisebox{1.5pt}{$\centerdot$} nodes assigned as flow end-points by sampling from inherited src/dst distributions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Parametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} session start time assigned according to Poisson arrival model \\
\raisebox{1.5pt}{$\centerdot$} nodes assigned as flow end-points by sampling from fitted BiPareto model
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Uniform}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} session start times selected uniformly from viable range \\
\raisebox{1.5pt}{$\centerdot$} nodes have the mean number of generic flows
\vspace{2pt}
\end{minipage} \\
\hline
\multirow{14}{*}[2.5pt]{\class{Flow}}
& \multirow{1}{*}[-0.05em]{\class{Trace}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} flow start time and size taken from trace \\
\raisebox{1.5pt}{$\centerdot$} each flow corresponds to a specific trace flow with an associated packet trace \\
\raisebox{1.5pt}{$\centerdot$} overrides distributions of size/interval packet behaviors with flow-specific versions \\
\raisebox{1.5pt}{$\centerdot$} packet stream from trace file
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Nonparametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} start time relative to previous flow sampled from inherited distribution \\
\raisebox{1.5pt}{$\centerdot$} flow size sampled from inherited distribution of flow sizes \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node \\
\raisebox{1.5pt}{$\centerdot$} generic packets generated until flow size exhausted
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Parametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} start time relative to previous flow sampled from fitted Lognormal model \\
\raisebox{1.5pt}{$\centerdot$} flow size sampled from fitted BiPareto model \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node \\
\raisebox{1.5pt}{$\centerdot$} generic packets generated until flow size exhausted
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Uniform}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} start time relative to previous flow equals mean inherited inter-flow interval \\
\raisebox{1.5pt}{$\centerdot$} all flows have size equal to inherited mean flow size \\
\raisebox{1.5pt}{$\centerdot$} inherits distributions of size/interval packet behaviors from destination node \\
\raisebox{1.5pt}{$\centerdot$} generic packets generated until flow size exhausted
\vspace{2pt}
\end{minipage} \\
\hline
\multirow{5}{*}[2.5pt]{\class{Packet}}
& \multirow{1}{*}[-0.05em]{\class{Trace}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} each packet corresponds to a specific trace packet
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Nonparametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals sampled from inherited distributions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Parametric}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals drawn from fitted Exponential distributions
\vspace{2pt}
\end{minipage} \\
\cline{2-3}
& \multirow{1}{*}[-0.05em]{\class{Uniform}} &
\begin{minipage}[l]{4.3in}
\vspace{2pt}
\raisebox{1.5pt}{$\centerdot$} all packet sizes and inter-packet intervals are equal to inherited mean
\vspace{2pt}
\end{minipage} \\
\hline
\end{tabular}