### 2008-06-13 tables/traffic_models.tex
Lots of work on Common Assumptions section.

%!TEX root = ../paper.tex

\begin{table*}
\begin{center}
\small
\begin{tabular}{|c|c|l|}
\multicolumn{1}{c}{\textbf{model}} &
\multicolumn{1}{c}{\textbf{transformation}} &
\multicolumn{1}{c}{\textbf{description}} \\
\hline
\caps{GMM} &
--- &
the general matrix model
\\\hline\hline
time uniform &
$\mat{A}'=\mat{A}(\ones/d_t)$ &
all start times are equally likely
\\\hline
node uniform &
$\mat{S}'=\mat{S}(\ones/n)$~~~~$\mat{D}'=\mat{D}(\ones/n)$ &
nodes behave stochastically identically with respect to flows
\\\hline
full uniform &
\textit{all uniform transformations} &
applies all the uniformity assumptions of the above models
\\\hline\hline
time marginal &
$\mat{A}'=(\ones/f)\mat{A}$ &
start times chosen from marginal distribution of start times
\\\hline
size marginal &
$\mat{B}'=(\ones/f)\mat{B}$ &
flow size (bytes) is chosen independently of other properties
\\\hline
node marginal &
$\mat{S}'=(\ones/f)\mat{S}$~~~~$\mat{D}'=(\ones/f)\mat{D}$ &
nodes behave distinctly, but behavior is independent of flows
\\\hline
packet marginal &
$\mat{Z}'=(\ones/f)\mat{Z}$~~~~$\mat{V}'=(\ones/f)\mat{V}$ &
packet behavior is variable bit-rate, aggregated across all flows
\\\hline
full marginal &
\textit{all marginal transformations} &
applies all the marginality assumptions of the above four models
\\\hline\hline
source conditional &
$\mat{F}'=(\trans{\mat{SS}}/f)\mat{F}$ &
flow behaviors are aggregated on a per source-node basis
\\\hline
dest conditional &
$\mat{F}'=(\trans{\mat{DD}}/f)\mat{F}$ &
flow behaviors are aggregated on a per destination-node basis
\\\hline
mixed conditional &
$\mat{F}'=\frac{1}{2f}(\trans{\mat{SS}}+\trans{\mat{DD}})\mat{F}$ &
flow behaviors are aggregated on a mixed source/destination basis
\\\hline
\end{tabular}
\caption{Matrix-based traffic models evaluated by paired differential simulation.}
\label{tab:traffic-models}
\end{center}
\vspace{-2em}
\end{table*}