### 2008-06-25 letter.tex
Letter addressing reviewer comments.

\documentclass{letter}

\oddsidemargin=.2in
\evensidemargin=.2in
\textwidth=5.9in
\topmargin=-.5in
\textheight=9in

\usepackage{cite,amsmath,amssymb,relsize,paralist,mathtools,units}

\input{definitions}

\newenvironment{thebibliography}[1]{%
\section{References}%
\begin{enumerate}%
}{%
\end{enumerate}%
}

\name{%
Stefan~Karpinski\\
Elizabeth~M.~Belding\\
Kevin~C.~Almeroth\\
John~R.~Gilbert%
}

\begin{document}
\begin{letter}{Reviewers of \caps{ACM}/Monet Manuscript \#MONE-95}

\opening{Dear Reviewers,}

Thank you for your thoughtful and helpful reviews of out \caps{ACM}/Monet submission.
From your responses and feedback from other peers, we believe that the first version of our paper did not adequately make clear the motivation, purpose or contribution of our work.
Therefore, we have significantly revised the introductory and motivational material of our paper as well as presentation of the technical concepts regarding the linear representation of network traffic as vectors and matrices.
The simulation results that support our claim that the general matrix model captures enough of the behavior of real traffic to be sufficiently realistic as a basis for generative models of network traffic remains the same except for the \model{mixed conditional} model (previously called \model{mixed regular}).
There was an error in the code generating traffic for this model, which caused the simulation results to be incorrect.
We have corrected this error.
In the new version, we have omitted the discussion of measures of error and statistical tests, which already appeared in our previous work~\cite{Karpinski07:cbr-failure}; it was deemed too far from the central topic of the paper.
Since the utility and significance of the general matrix model was not made clear in our previous version of the paper, we have added in the discussion section preliminary results from using the numerical linear technique of nonnegative matrix factorization on the \caps{GMM} matrices describing the packet behaviors of flows.
We believe these results are very exciting and promising, and indicate the utility and potential of the representing network traffic linearly.
In what follows, we indicate where in the paper revision, we address specific comments, questions and concerns raised by our collective reviewers.

\begin{enumerate}

\item \textbf{Information destruction in uniform, marginal and conditional (n\'e regular) models, seen through linear representation.}
It was mentioned that it was unclear how the transformations allow to understand the effect and drawbacks of each simplified uniform, marginal or conditional model.
In the revised paper, we have observed that the uniform, marginal and conditional transformations are all multiplications by specific low-rank matrices.
Since low-rank matrices are by definition uninvertible, the information content of the original traffic matrix is irrevocably destroyed by these transformations.
The revised paper discusses this issue in several places:
\begin{enumerate}
\item Introduction, page 2, column 2, paragraph 2;
% TODO: add another paragraph about this?
\item Discussion, page 15, column 2, paragraph 1.
\end{enumerate}
In the discussion section, moreover, we propose matrix factorization, rather matrix multiplication as a less destructive and thus more faithful method of reducing the complexity, size and dimension of traffic representations.

\item \textbf{Computational complexity of \caps{GMM} compared with simplified models.}
% TODO: address complexity in discussion.
This issue is now addressed in our discussion section.
We admit that large matrix representations are more computationally costly than simpler models.
The purpose of this work, however, is to establish how traffic can be represented and generated realistically at all, not how it can be done most efficiently.
Currently, realistic traffic modeling is an unsolved problem.
We observe, however, that generating application traffic for a simulation takes orders of magnitude less \caps{CPU}-time than running the simulation itself.
Therefore, traffic model complexity is considered a non-issue.

\item \textbf{Comprehensiveness of performance metrics used in evaluation.}
% TODO: address comprehensiveness of metrics in discussion.
We already briefly addressed this issue in the original version of the paper:
we have evaluated a variety of other performance metrics, but omit them here due to lack of space.
Moreover, the performance comparisons for the metrics presented here are indicative of other performance metrics.
We address this now in our discussion section in greater detail.

\item \textbf{Miscellaneous other questions and concerns.}
% TODO: make sure this is addressed adequately.
We have addressed the inter-relation between flow size, packet behavior, duration and start time in the section describing traffic generation (page 10, section 4.4).
The adequacy of the 24-hour trace is discussed in the discussion section: in short, 24 hours of data is not enough to draw general conclusions about all traffic, but is enough to demonstrate the potential effectiveness, or provide counter-examples to the effectiveness of modeling techniques; this is all we attempt to do here.
Further validation of methods on larger sets of data is required in the future.
We provide a concrete example of theoretical and numerical tools and their application to the \caps{GMM} in the use of nonnegative matrix factorization to understand and explain flow packet behaviors.
This technique and preliminary results from its application are shown in the discussion section.
The seven properties used in the \caps{GMM} are 
We note in our revised related work section that the work of \FHC~\cite{Hernandez06:dissertation} perfectly complements ours in that it provides the means to emulate the closed-loop behavior of applications using \caps{TCP}, but does not provide any models of application behavior;
our work provides the necessary models of applications behavior.
See Related Work, page 3, column 2, paragraph 2 for our discussion of this issue.

\end{enumerate}

We hope that the revised version addresses your concerns and questions adequately and that it meets your expectations for professional scholarly publications.

\closing{Yours Respectfully,}
\signature{Stefan~Karpinski}

\vfill
\bibliographystyle{plain}
\bibliography{references}

\end{letter}

\end{document}