### 2008-11-04 tables/traffic_models.tex
Corrections to the proof text.

%!TEX root = ../paper.tex

\begin{table*}
\begin{center}
\small
\begin{tabular}{|c|c|l|}
\multicolumn{1}{c}{\textbf{model}} &
\multicolumn{1}{c}{\textbf{transformation}} &
\multicolumn{1}{c}{\textbf{description}} \\
\hline
\caps{GMM} &
--- &
the general matrix model
\\\hline\hline
time uniform &
$\mat{A}' = (\nicefrac{1}{n})\,\mat{A}\ones$ &
all start times are equally likely
\\\hline
node uniform &
$\mat{S}' = (\nicefrac{1}{n})\,\mat{S}\ones \quad
 \mat{D}' = (\nicefrac{1}{n})\,\mat{D}\ones$ &
all nodes are equally likely as source or destination
\\\hline
full uniform &
\textit{all uniform transformations} &
all start times, sources and destinations are equally likely
\\\hline\hline
time marginal &
$\mat{A}' = (\nicefrac{1}{f})\,\ones\mat{A}$ &
start times chosen from marginal distribution of start times
\\\hline
size marginal &
$\mat{B}' = (\nicefrac{1}{f})\,\ones\mat{B}$ &
flow size (bytes) is chosen from marginal distribution of sizes
\\\hline
node marginal &
$\mat{S}' = (\nicefrac{1}{f})\,\ones\mat{S} \quad
 \mat{D}' = (\nicefrac{1}{f})\,\ones\mat{D}$ &
source and dest. chosen marginally, independent of other properties
\\\hline
packet marginal &
$\mat{Z}' = (\nicefrac{1}{f})\,\ones\mat{Z} \quad
 \mat{V}' = (\nicefrac{1}{f})\,\ones\mat{V}$ &
packet behavior is variable bit-rate, marginal across all flows
\\\hline
full marginal &
\textit{all marginal transformations} &
applies all the marginality assumptions of the above four models
\\\hline\hline
source conditional &
$\mat{F}' = \rStoch(\trans{\mat{S}}\mat{S})\,\mat{F}$ &
flow behaviors are aggregated on a per source-node basis
\\\hline
dest. conditional &
$\mat{F}' = \rStoch(\trans{\mat{D}}\mat{D})\,\mat{F}$ &
flow behaviors are aggregated on a per destination-node basis
\\\hline
% mixed conditional &
% $\mat{F}' = \nicefrac{1}{2}\,(
%               \rStoch(\trans{\mat{S}}\mat{S}) +
%               \rStoch(\trans{\mat{D}}\mat{D})
%             )\,\mat{F}$ &
% flow behaviors are aggregated on a mixed source/destination basis
% \\\hline
\end{tabular}
\caption{Matrix-based traffic models used in paired differential simulation experiments.}
\label{tab:traffic-models}
\end{center}
\vspace{-1em}
\end{table*}